---
title: "problem7-2"
author: "Mehrab Atighi"
date: "11/19/2021"
output: 
  beamer_presentation:
    fig_width: 9
    fig_height: 6
    theme: Warsaw
---

## problem
\footnotesize
\textbf{Personal Loan Acceptance}. Universal Bank is a relatively young bank growing
rapidly in terms of overall customer acquisition. The majority of these customers are
liability customers (depositors) with varying sizes of relationship with the bank. The
customer base of asset customers (borrowers) is quite small, and the bank is interested
in expanding this base rapidly to bring in more loan business. In particular, it wants
to explore ways of converting its liability customers to personal loan customers (while
retaining them as depositors).\newline
A campaign that the bank ran last year for liability customers showed a healthy
conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use
k-NN to predict whether a new customer will accept a loan offer. This will serve as
the basis for the design of a new campaign.\newline
The file UniversalBank.csv contains data on 5000 customers. The data include
customer demographic information (age, income, etc.), the customer’s relationship
with the bank (mortgage, securities account, etc.), and the customer response to the
last personal loan campaign (Personal Loan). Among these 5000 customers, only 480
(= 9.6%) accepted the personal loan that was offered to them in the earlier campaign.
Partition the data into training (60%) and validation (40%) sets.

## problem

a. Consider the following customer:
Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1
= 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0,
CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification
with all predictors except ID and ZIP code using k = 1. Remember to transform
categorical predictors with more than two categories into dummy variables first.
Specify the success class as 1 (loan acceptance), and use the default cutoff value of
0.5. How would this customer be classified?

b. What is a choice of k that balances between overfitting and ignoring the predictor
information?

c. Show the confusion matrix for the validation data that results from using the best k.

## problem 

d. Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.

e. Repartition the data, this time into training, validation, and test sets (50% : 30% :
20%). Apply the k-NN method with the k chosen above. Compare the confusion
matrix of the test set with that of the training and validation sets. Comment on the
differences and their reason.

## solution 
\tiny
At the first we want to introduction our dataset in R.
```{r ,echo=TRUE}
Data<-read.csv("F:/lessons/Data mining/Data/UniversalBank.csv")
dim(Data)
head(Data,4)
```

## solution
\footnotesize
Now we want to transform Educatin variable to tree dummy variables, and we names them :\newline
Education_1 , Education_2 , Educatioin_3 .
```{r ,echo=TRUE}
data=Data[,-c(1,5)]
data$Education_1=c(1:nrow(data))
data$Education_2=c(1:nrow(data))
data$Education_3=c(1:nrow(data))
for(i in 1:nrow(data)){
  if(data$Education[i]==1){
    data$Education_1[i] = 1 
  }else(data$Education_1[i] = 0)
  if(data$Education[i]==2){
    data$Education_2[i] = 1 
  }else(data$Education_2[i] = 0)
  if(data$Education[i]==3){
    data$Education_3[i] = 1 
  }else(data$Education_3[i] = 0)
}
data$Education=c()
```
\normalsize

## solution
\tiny
Now we want to see new data heads and dimantion.
```{r ,echo=TRUE}
head(data)
dim(data)
summary(data)
```
\normalsize

## solution
we want to select  the train and test subsets of dataset.
```{r , echo=TRUE}
set.seed(5)
n=sample(c(0,1) , nrow(data) ,
         prob = c(0.6 , 0.4) ,replace = TRUE)
train_data=data[which(n==0),]
test_data=data[which(n==1),]

```

## solution
\tiny
we are going to work with normalized data so we should make normal data frame, and using preProcess function to normalizing variables.
```{r ,echo=TRUE}
train_normal<- train_data
test_normal<-test_data
data_normal<-data
new.df=data.frame(Age = 40, Experience = 10 , Income = 84,
                  Family = 2 , CCAvg = 2, Education_1= 0 ,Education_2= 1,
                  Education_3=0 , Mortgage = 0 , Securities.Account = 0 ,
                  CD.Account = 0 , Online = 1 , CreditCard =1)

#install.packages("caret")
library(caret)

norm.values<-preProcess(train_data[,-7] , method = c("center" , "scale"))
train_normal<-predict(norm.values , train_data[,-7])
test_normal<-predict(norm.values , test_data[,-7])
data_normal<-predict(norm.values , data[,-7])
new.norm.df <-predict(norm.values, new.df)
n2=sample(c(0,1) , nrow(train_normal) ,
         prob = c(0.6 , 0.4) ,replace = TRUE)
train_normal_2=train_normal[which(n2==0),]
train_test_normal=train_normal[which(n2==1),]
```
\normalsize

## solution
\tiny
Now we want to select K value, we should work with all of data
```{r ,echo=TRUE ,warning=FALSE}
#b)
memory.limit(size = 9999999999999)
library(caret);library(e1071)
accuracy_data=data.frame(k = seq(1,nrow(train_normal_2),1) , accuracy=rep(0,nrow(train_normal_2)))
system.time(
  for(i in 1:nrow(accuracy_data)){
    knn.predict<-FNN::knn(train =train_normal_2[,-7] ,test = train_test_normal[,-7] ,
                          cl = train_normal_2[,7] ,k=i)
    accuracy_data[i , 2] = confusionMatrix(data = knn.predict[1:nrow(train_test_normal)],
                                           factor(train_test_normal[,7]) ,)$overall[1]
  })
head(accuracy_data)
(best = which.max(accuracy_data$accuracy))
```

## solution
\small
```{r ,echo=TRUE}
#c)
knn.predict<-FNN::knn(train =train_normal[,-7] ,
                      test = test_normal[,-7] ,
                      cl = train_data[,7] ,k=best)
confusionMatrix(data = knn.predict[1:nrow(test_data)],
                factor(test_data[,7]) )$overall[1]
```

## solution
\small
```{r , echo=TRUE}
#d)
library(FNN)
new.df=data.frame(Age = 40, Experience = 10 , Income = 84,
                  Family = 2 , CCAvg = 2, Education_1= 0 ,
                  Education_2= 1,Education_3=0,Mortgage = 0,
                  Securities.Account = 0 ,CD.Account = 0 ,
                  Online = 1 , CreditCard =1)

nn<-knn(train = train_normal , test = new.df,
        cl =train_data[,7] ,k=best)
r=row.names(train_data)[attr(nn, 'nn.index')]
train_data[r,7]
```
\normalsize


## solution
\footnotesize
```{r ,echo=TRUE}

#e)
n=sample(c(0,1,2) , nrow(data) ,prob = c(0.5 , 0.3 ,0.2) ,replace = TRUE)
train_data_2=data[which(n==0),]
valid_data=data[which(n==1),]
test_data_2=data[which(n==2),]

norm.values<-preProcess(train_data_2[,-7] , method = c("center" , "scale"))
train_normal_2<-predict(norm.values , train_data_2[,-7])
test_normal_2<-predict(norm.values , test_data_2[,-7])
valid_normal<-predict(norm.values , valid_data[,-7])
data_normal_2<-predict(norm.values , data[,-7])
```
\normalsize

## solution
\tiny
```{r , echo=TRUE}
#for test data:
knn.predict<-FNN::knn(train = train_normal_2[,-7] ,test = test_normal_2[,-7] ,
                      cl = train_data_2[,7] ,k=best)
confusionMatrix(data = knn.predict[1:nrow(test_data_2)],
                factor(test_data_2[,7]) )$overall[1]
#for valid data:
knn.predict<-FNN::knn(train =train_normal_2[,-7] ,test = valid_normal[,-7] ,
                      cl = train_data_2[,7] ,k=best)
confusionMatrix(data = knn.predict[1:nrow(valid_normal)],
                factor(valid_data[,7]) )$overall[1]
#for train data:
knn.predict<-FNN::knn(train =train_normal_2[,-7] ,test = train_normal_2[,-7] ,
                      cl = train_data_2[,7] ,k=best)
confusionMatrix(data = knn.predict[1:nrow(train_normal_2)],
                factor(train_data_2[,7]) )$overall[1]
```

