---
title: "Principal Components Analysis in R for thired exercise"
author: "Mehrab Atighi - Zahra Ghasabi - Zahra Barzegar - Morteza Hoseingholipur -Mina Daliry"
date: "10/19/2021"
output: 
  beamer_presentation: 
    fig_width: 9
    fig_height: 6
    theme: Warsaw
---

## Contents
- Whats the issue?\newline
- How to use PCA in R \newline
- How to visualizating PCA Data in R \newline



## Whats the issue?


We want to display two way to dimantion reduction our data in R.\newline
When we want to do Dimantion reduction we should make a data as type of \textcolor{blue}{Data Frame}, after that we should use \textcolor{orange}{princomp} function that is in base of R.\newline
For Example we have \textcolor{blue}{ramus data(exercise3)}. we import Dataset from Excel csv file with \textcolor{orange}{read.csv("addres",...)} into R. and using \textcolor{orange}{head(data name ,number to show)} to see the 4 observation of our data. and using \textcolor{orange}{dim} to see the dimantion of our data.

## Whats the issue?

\tiny
```{r warning=FALSE}
data<-read.csv("F:/lessons/Multi countios Variate2/data/ramos-data.csv")
data=data[,-1]
#View(data)
head(data , 4)
dim(data)
eigen(cor(data))
```

## Principal Components Analysis in R


Now we are going to use Principal Components Analysis with auto way in R:\newline
we should use \textcolor{orange}{princomp} function.\newline
we should write the data as type as \textcolor{blue}{data.frame} into this function.\newline
we have a \textcolor{orange}{cor} argument in this function that we can equal it with \textcolor{blue}{TRUE} when we want to do Principal Components method with Correlation matrix, or \textcolor{blue}{FALSE} when we want to do Principal Components method with Covariance matrix.\newline
the other argument in this function is \textcolor{orange}{scores} that we can equal it with \textcolor{blue}{TRUE} and it will show the score of each component.

```{r warning=FALSE}
pc.r<-princomp( data,cor = TRUE ,scores = TRUE )
pc.c<-princomp(data , cor = FALSE ,scores = TRUE)
```
\normalsize

## Principal Components Analysis in R
\small
```{r warning=FALSE}
summary(pc.r)
```
\normalsize
```{r warning=FALSE}

head(pc.r$loadings,2)
head(pc.r$scores, 2)
```

## Principal Components Analysis in R
\small
```{r warning=FALSE}
summary(pc.c)
```
\normalsize
```{r warning=FALSE}
head(pc.c$loading,2)
head(pc.c$scores, 2)
```

## Principal Components Analysis in R

to see the Percentage of Explained Variances of each variable we should install a new packages that we call \textcolor{blue}{factoextra} and library it.\newline
the output is a scree plot that we can see codes in next silde.

## How to visualizating PCA Data in R


\tiny
```{r warning=FALSE, out.height="65%", out.width="85%"}
#install.packages(factoextra)
library(factoextra)
fviz_eig(pc.c)#plot(pc.c)
#fviz_eig(pc.r)#plot(pc.r)
```

## How to visualizating PCA Data in R
\small
So Now we select 2 components of our data here and ploting them.\newline
we can use 2 function here, the first function is plotting components 1 , 2 with \textcolor{orange}{plot()} function and the second function is \textcolor{orange}{biplot}.
```{r, out.height="65%", out.width="85%"}
plot(pc.r$scores[,1],pc.r$scores[,2]
     ,xlab = "Comp1" , ylab="Comp2" ,col="Blue")
#biplot(pc.r)
```
\normalsize

## How to visualizating PCA Data in R
\tiny
Now we want to see the Graph of individuals. Individuals with a similar profile are grouped together.
```{r, out.height="65%", out.width="85%"}
 #biplot(pc.r)
 fviz_pca_ind(pc.r,
              col.ind = "cos2", # Color by the quality of representation
              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
              repel = TRUE)      #Avoid text overlapping)
```
\normalsize

## How to visualizating PCA Data in R

```{r, out.height="65%", out.width="85%"}
fviz_pca_biplot(pc.r, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```
End.


