---
title: "final test"
author: "Mehrab Atighi"
date: "6/20/2021"
output: word_document
---

```{r}
set.seed(1)
library("ISLR")
data("swiss")
attach(swiss)
head(swiss)
```

# a)

## for make Full model and plot we have:

```{r}
fit<-lm(Fertility~. , data=swiss)
summary(fit)
par(mfrow=c(2,2))
plot(fit)
```


we can see that just Examination and Agriculture variables (predictors) 
are not signifact and anothers are becuse they have a p-values less than
0.05(alpha). 
the adjust R sqrue is 0.67  and R squre is 0.7 and our standard error is 7.165
and Agriculture , Examination , Education predictors have
negative relationships with our response.

## for make plot of each varibale with response we have:

```{r}
par(mfrow=c(1,1))
plot(Fertility,Agriculture , col="red")
abline(lm(Fertility~Agriculture), col="Blue")
```


According to this plot we see the approxmiatly posetive realationships
and we have 1 High leverage point (the left and down side of plot).

```{r}
plot(Fertility,Examination , col="red")
abline(lm(Fertility~Examination), col="Blue")
```


now we can see the negative realationships between response and this predictor
and we have 3 High leverage points (the left side).

```{r}
plot(Fertility,Education , col="red")
abline(lm(Fertility~Education), col="Blue")
```


now we can see the negative realationships between response and this predictor
and we have 3 High leverage points (the left side).

```{r}
plot(Fertility,Catholic , col="red")
abline(lm(Fertility~Catholic), col="Blue")
```


we can see the posetive realationships and 3High leverage points (the left side).

```{r}
plot(Fertility,Infant.Mortality , col="red")
abline(lm(Fertility~Infant.Mortality), col="Blue")
```

 
 here we can see the posetive realationships with one outlires and 
 high leverage point.

## for make covariance matrix of variables we have:
```{r}
cov(swiss)
```

according to this out put we see all of the covariance between each variables.

# b)

## best subset selection:
at the first we need to library this package:

```{r}
library(leaps)
```

Now we make our model from swiss data:
```{r}
best.subset.fit1<-regsubsets(Fertility~. , data = swiss , nvmax = 19)
summary(best.subset.fit1)
```


now here we can see that in the first step we have select M1 model or
a model with one predictors from all of the models with one predictors 
and we can see that this method choose the Education predictors from these models.
like this for the model with 2 predictors we choose model with M1 + Catholic=M2
predictors from all of them.
like this for the model with 3 predictors we choose model with  M2+Infant.Mortality =M3
predictors from all of them.
like this for the model with 4 preditos we choose model with M3+Agriculture =M4
and at the last for model with all of predictors we have just 1 model
or full model with all of the predictors M4 +Examinarion=M5


# backward subset selection:
```{r}
backward.fit1<-regsubsets(Fertility~.,data=swiss ,nvmax= 19 , method = "backward")
summary(backward.fit1)
```


here at the first we have compelte model(M0) with all of the predictors
in the M1 we just delete the Education preditor from our M0 it means that M1=M0-Education
in the next step we have: M2= M1-Catholic
the next step we have: M3= M2-Infant.mortality
the next step we have: M4 = M3 - Agriculture
the M5 is the model with 0 predictor. M5= M4-Examination


# c)


Now we want to make a training and test data with probabilty (0.7 , 0.3 ) and again chek models with bic:

```{r}
sample<-sample(c(TRUE , FALSE ) , nrow(swiss) , replace = T , prob=c(0.7,0.3))
train<-swiss[sample,]
test<-swiss[!sample,]
```


Now again we make model with best subset selection and backward method here:
```{r}
best.subset.fit2<-regsubsets(Fertility~. , data = train , nvmax = 19)
backward.fit2<-regsubsets(Fertility~.,data= train ,nvmax= 19 , method = "backward")
```


now we want to calculate and see the bic of each method here:
```{r}  
bic1<-summary(best.subset.fit2)$bic
plot(bic1 , type="b" , col="Blue" , xlab="number of predictors")
points(4,bic1[which.min(bic1)] ,cex=2 , col="red")
```


according to this plot we choose the model with 4 predictors that have the minimum of the BIC.

```{r}
bic2<-summary(backward.fit2)$bic
plot(bic2 , type="b" , col="Blue",xlab="number of deleted predictors")
points(4,bic2[which.min(bic1)] ,cex=2 , col="red")
```


according to this plot we choose the model with 1 predictor that have the minimum of the BIC.attention:backward method.


End.
by: Mehrab Atighi