---
title: "IranKhodro timeseries Analysis"
author: "Mehrab Atighi"
date: "12/20/2021"
output: 
  beamer_presentation: 
    theme: Warsaw
---


                     In The Name Of God
                     
                     
## Contents
- problem statement \ref{1.0} \newline
- Dawnloading Data \ref{2.0} \newline
- Importing Data To R \ref{3.0} \newline
- PreProccesing of data \ref{4.0} \newline
- Regression modeling for Data \ref{5.1} \newline
- Selecting and Ploting Data \ref{5.0} \newline
- plot of auto correlation function \ref{6.0} \newline
- plot of  partial auto correlation function \ref{7.0} \newline
- Cheking The ARIMA Models with Data \ref{8.0} \newline
- Modeling for all states \ref{9.0} \newline

## problem statement
\label{1.0}
We want to use the Timeseries Algorithms On the shares of Irankhodro Company in the Iranian stock market.\newline
we should chek it with all of \textcolor{blue}{ARIMA} state until the second time.\newline

## Dawnloading Data
\label{2.0}
So We went Tse website and dawnload needed Excel file.for this step we should go to the \href{http://www.tsetmc.com/Loader.aspx?ParTree=151311&i=65883838195688438}{Site} and after dawnloding Excel file, we should save that as \textcolor{blue}{CSV} format.\newline

## Importing Data To R 
\label{3.0}
\small
Now we want to import the Dataset in to R. so we should use \textcolor{orange}{read.csv("...")} function.
```{r , echo=TRUE , warning=FALSE}
df = read.csv(file = "F:/lessons/time series/project/Iran.Khodro.csv")
head(df , 4)
```

## PreProccesing of data
\label{4.0}
\tiny
we should separate our data and we just need to Date and close value for each day.
so we use \textcolor{orange}{separate} function from \textcolor{blue}{tidyverse} Package.\newline
```{r ,echo=TRUE , warning=FALSE}
library(tidyverse)
df= separate(df,2,c("year","mouth","day"),
           sep =c(4,6),remove = TRUE)
df= df[,c(2,3,4,8)]
head(df , 4 )
```
\normalsize


## Regression modeling for Data 
\label{5.1}
\tiny
Now we want to make a regression model for our data and calculating the coefficients estimate.

```{r ,echo=TRUE , warning=FALSE}
response = df$X.CLOSE.[(nrow(df)-200):nrow(df)]
model_reg = lm(response ~ (time(response)))
summary(model_reg)
```
\normalsize

## Selecting and Ploting Data
\label{5.0}
\small
Now we want to select the Two hundred days ago data to timeseries alg and ploting them fron dataset.
```{r ,echo=TRUE , warning=FALSE ,out.width="70%" ,out.height="70%"}
plot(response,type="o" , col = "orange")
abline(model_reg)
```
\normalsize

## plot of auto correlation function
\label{6.0}
Now we want to ploting the auto correlation function values of dataset.
```{r ,echo=TRUE , warning=FALSE,out.width="70%" ,out.height="70%"}
acf(response,lag=50,ylab="acf")
```

## plot of  partial auto correlation function
\label{7.0}
We want to ploting the partial auto correlation function values of dataset.
```{r ,echo=TRUE , warning=FALSE,out.width="60%" ,out.height="60%"}
pacf(response,lag=50,ylab="pacf")
```
We can see that the pacf values from the first times is lower than a blue bound and we can say its approximatly zero.

## Cheking The ARIMA Models with Data
\label{8.0}
Now we want to chek all of ARIMA state until two times.\newline
- ARIMA(0,0,1)\ref{8.1}
- ARIMA(0,0,2)\ref{8.2}
- ARIMA(1,0,0)\ref{8.3}
- ARIMA(2,0,0)\ref{8.4}
- ARIMA(1,0,1)\ref{8.5}
- ARIMA(2,0,2)\ref{8.6}
- ARIMA(0,1,1)\ref{8.7}
- ARIMA(0,1,2)\ref{8.8}
- ARIMA(1,1,0)\ref{8.9}
- ARIMA(2,1,0)\ref{8.10}
- ARIMA(1,1,1)\ref{8.11}
- ARIMA(2,1,2)\ref{8.12}
- ARIMA(0,2,1)\ref{8.13}
- ARIMA(0,2,2)\ref{8.14}
- ARIMA(1,2,0)\ref{8.15}
- ARIMA(2,2,0)\ref{8.16}
- ARIMA(1,2,1)\ref{8.17}
- ARIMA(2,2,2)\ref{8.18}

## Modeling for all states
\label{9.0}
\tiny
We Want to Chek the response with ARIMA(0,0,1)\label{8.1}:
attention that we need to install and library the \textcolor{blue}{TSA} Package.
```{r ,echo=TRUE , warning=FALSE}
library(TSA)
(M1=arima(response,order=c(0,0,1)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(0,0,1)\label{8.2}:
```{r ,echo=TRUE , warning=FALSE }
(M2=arima(response,order=c(0,0,2)))
```
\normalsize


## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(1,0,0)\label{8.3}:
```{r ,echo=TRUE , warning=FALSE }
(M3=arima(response,order=c(1,0,0)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(2,0,0)\label{8.4}:
```{r ,echo=TRUE , warning=FALSE }
(M4=arima(response,order=c(2,0,0)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(1,0,1)\label{8.5}:
```{r ,echo=TRUE , warning=FALSE }
(M5=arima(response,order=c(1,0,1)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(2,0,2)\label{8.6}:
```{r ,echo=TRUE , warning=FALSE }
(M6=arima(response,order=c(2,0,2)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(0,1,1)\label{8.7}:
```{r ,echo=TRUE , warning=FALSE }
(M7=arima(response,order=c(0,1,1)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(0,1,2)\label{8.8}:
```{r ,echo=TRUE , warning=FALSE }
(M8=arima(response,order=c(0,1,2)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(1,1,0)\label{8.9}:
```{r ,echo=TRUE , warning=FALSE }
(M9=arima(response,order=c(1,1,0)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(2,1,0)\label{8.10}:
```{r ,echo=TRUE , warning=FALSE }
(M10=arima(response,order=c(2,1,0)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(1,1,1)\label{8.11}:
```{r ,echo=TRUE , warning=FALSE }
(M11=arima(response,order=c(1,1,1)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(2,1,2)\label{8.12}:
```{r ,echo=TRUE , warning=FALSE }
(M12=arima(response,order=c(2,1,2)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(0,2,1)\label{8.13}:
```{r ,echo=TRUE , warning=FALSE }
(M13=arima(response,order=c(0,2,1)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(0,2,2)\label{8.14}:
```{r ,echo=TRUE , warning=FALSE }
(M14=arima(response,order=c(0,2,2)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(1,2,0)\label{8.15}:
```{r ,echo=TRUE , warning=FALSE }
(M15=arima(response,order=c(1,2,0)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(2,2,0)\label{8.16}:
```{r ,echo=TRUE , warning=FALSE }
(M16=arima(response,order=c(2,2,0)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(1,2,1)\label{8.17}:
```{r ,echo=TRUE , warning=FALSE }
(M17=arima(response,order=c(1,2,1)))
```
\normalsize

## Modeling for all states
\scriptsize
We Want to Chek the response with ARIMA(2,2,2)\label{8.18}:
```{r ,echo=TRUE , warning=FALSE }
(M18=arima(response,order=c(2,2,2)))
```
\normalsize

## Final AIC Comparing

Now we want to compare all models AIC .

```{r ,echo=TRUE , warning=FALSE}

(AIC=c(M1$aic,M2$aic,M3$aic,M4$aic,M5$aic,M6$aic,
  M7$aic,M8$aic,M9$aic,M10$aic,M11$aic,M12$aic,
  M13$aic,M14$aic,M15$aic,M16$aic,M17$aic,M18$aic))

min(AIC)
which.min(AIC)
```

So we can say that the M14 or ARIMA(0,2,2) is the best fit for our dataset.



## Final predict Visual

```{r ,echo=TRUE}
plot(response,type="o" , col = "orange")
with(df , lines(x = time(response) , y = fitted(M14)))
```

