---
title: "Exercise section 19"
author: "Mehrab Atighi"
date: "5/17/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#solve 
at the first we should inputs our data to R:
note: We have three response and 3 predictor.

```{r}
y1<-c(41.5,33.8,27.7,21.7,19.9,15.0,12.2,4.3,19.3,6.4,37.6,18,26.3,9.9,25,14.1,15.2,
      15.9,19.6)
y2<-c(45.9,53.3,57.5,58.8,60.6,58,58.6,52.4,56.9,55.4,46.9,57.3,55,58.9,50.3,61.1,
      62.9,60,60.9)
y3<-c(11.2,11.2,12.7,16,16.2,22.6,24.5,38,21.3,30.8,14.7,22.2,18.3,28.0,22.1,23.0,
      20.7,22.1,19.3)

x1<-c(162,162,162,162,172,172,172,172,167,177,157,167,167,167,167,177,177,160,160)     
x2<-c(23,23,30,30,25,25,30,30,27.5,27.5,27.5,32.5,22.5,27.5,27.5,20,20,34,34)
x3<-c(3,8,5,8,5,8,5,8,6.5,6.5,6.5,6.5,6.5,9.5,3.5,6.5,6.5,7.5,7.5)

```

Now we should make a Multivariate Multiple linear Regression model:
```{r}
fit<-lm(cbind(y1,y2,y3)~x1+x2+x3)
summary(fit)
```
According to outputs of this model we can say that the predictors are signifact for y1 or first response Becuse p-Values are lower than alpha = 0.05
for the second response or y2 just x1 is signifact and good response and another predictors are not good(according to the p-value).
for the thired response or y3 we can say that just the x3 is not signifact predictor and another are signifact and good for our models.
at the end we can see the resudual standard error and R- squred value.


Now we want to see the coefficients of our model and chek them with Betha in our question.
```{r}
fit$coefficients
```


