---
title: "sec12"
author: "Mehrab Atighi"
date: "4/15/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## a) Introducing data to R

```{r }

y1<-c(47.8,46.4,46.3,45.1,47.6,52.5,51.2,49.8,48.1,45.0,51.2,48.5,52.,48.2,49.6,50.7,47.2,53.3,46.2,46.3)
y2<-c(48.8,47.3,46.8,45.3,48.5,53.2,53.0,50.0,50.8,47.0,51.4,49.2,52.8,48.9,50.4,51.7,47.7,54.6,47.5,47.6)
y3<-c(49.0,47.7,47.8,46.1,48.9,53.3,54.3,50.3,52.3,47.3,51.6,53.0,53.7,49.3,51.2,52.7,48.4,55.1,48.1,51.3)
y4<-c(49.7,48.4,48.5,47.2,49.3,53.7,54.5,52.7,54.4,48.3,51.9,55.5,55.0,49.8,51.8,53.3,49.5,55.3,48.4,51.8)

dat<-data.frame(y1,y2,y3,y4)
dat

```

 H0: mean(y1,y2,y3,y4) == mu
 
 H1: mean(y1,y2,y3,y4) != mu

## b) 

Now we want to test all variable with HotelingsT2 function

```{r echo=TRUE}
library("ICSNP")

HotellingsT2(dat,mu= c(48,49,50,51),test = "chi")

```
According to this test we can say that the H0 accept becuse our p-value is more than 0.05(alpha)
and it means that the lower jaw bone size of boys are not equal with mu(48,49,50,51) in 8,8.5,9,9.5 years old.


## c) 

Now we want to test each variable mean with mu value
```{r }
#H0: mean(y1) == 48
 
#H1: mean(y1) != 48

t.test(y1,mu = 48)
```

according to tests outputs we can say that the p-value is more that alpha and H0 accept and we can say the mean of y1 approximately is equal to 48.65

```{r}
#H0: mean(y2) == 49
 
#H1: mean(y2) != 49
t.test(y2, mu = 49)
```

according to tests outputs we can say that the p-value is more that alpha and H0 accept and we can say the mean of y2 approximately is equal to 49.625

```{r}
#H0: mean(y3) == 50
 
#H1: mean(y3) != 50
t.test(y3, mu = 50)
```

according to tests outputs we can say that the p-value is more that alpha and H0 accept and we can say the mean of y3 approximately is equal to 50.57

```{r}
#H0: mean(y4) == 51
 
#H1: mean(y4) != 51
t.test(y4, mu = 51)

```

according to tests outputs we can say that the p-value is more that alpha and H0 accept and we can say the mean of y4 approximately is equal to 51.45







