---
title: "Calculate Cv value with diffrent polynomial degree regression."
author: "Mehrab Atighi"
date: "5/4/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# library the packages that we need to solve this exercise
after the library the needed packages we should library our data 

```{r}
library(ISLR)
library(reshape2)
library(tibble)
library(ggplot2)
data(Auto)
```

Now we want to use leave one out cross validation to calculate Cv with diffrent polynomial degree regression.
At the first we can see summary of our data 

```{r}
# Leave-one-out cross-validation :
dt<-data.frame(Auto$mpg , Auto$horsepower)
head(dt)
```

Now we make a table and make loops for calculate each MSE with diffrent polynomial degree regression.

```{r}
all.MSE.table<-tibble(number.of.leave.out.data = rep(c(1:nrow(dt)),10) , degree = rep(c(1:10),each=nrow(dt)) ,MSE =c(1:3920))
# i show the degree of the polynomial regression function
# j show the number of data that we will leave out (MSE_j) for each i
test<-data.frame(dt$Auto.horsepower ,  dt$Auto.mpg)
count=1
for(i in 1:10){
  for(j in 1:nrow(dt)){
    train<-data.frame(train1=dt$Auto.mpg[-j] , train2=dt$Auto.horsepower[-j])
    fit<-glm( train1 ~poly(train2,degree= i),data = train)
    new= data.frame(train2 = test$dt.Auto.horsepower[j])
    all.MSE.table$MSE[count]<-(test$dt.Auto.mpg[j]-predict(fit ,newdata = new ,type="response"))^2
    count=count+1
    }
}

```

Now we can see the all of the MSE (mean squre Error) with diffrent polynomial degree regression.

```{r}
all.MSE.table
```

Now we want to calculate Cv with MSE values , then see them.

```{r}
Cv=c(1:10)
Cv.table<-tibble(Degree =1:10 ,Cv)

for(i in 1:10){
Cv.table$Cv[i]=mean(all.MSE.table$MSE[which(all.MSE.table$degree==i)])
}

Cv.table
```



# drawing plots
Now we want to draw the plot of y and y hat

```{r}
ggplot(all.MSE.table , aes(degree,MSE))+
  geom_point(colour="Red",cex=0.1)


```


Now we want to draw the plot of Cv

```{r}
ggplot(Cv.table,aes(Degree , Cv))+
  geom_point(colour = "Red",cex=4) +
  geom_line(colour = "Blue")
  


```


End.

