---
title: "Ploting model 3 with R"
author: "Mehrab Atighi"
date: "2022-10-11"
output: 
  beamer_presentation: 
    fig_width: 9
    fig_height: 6
    theme: Warsaw
---

## Contents

- problem  \ref{Problem} \newline
- Calculating Distribution function \ref{Dis} \newline
- ploting Cummulative distribution function \ref{CDF} \newline
- calculating Survial function \ref{Sur} \newline
- ploting Survial function \ref{Psur} \newline
- calculating probability mass function \ref{pdf} \newline
- ploting probability mass function \ref{Ppdf} \newline

## Problem 
\label{Problem} 

According to the \textcolor{pink}{Model 3  in chapter 2} of \textcolor{blue}{Loss Models from Data to Decision Edition 5}, We know that \newline
Model 3 This random variable could serve as a model for the number of claims on one policy in one year. Probability is concentrated at the five points (from 0 to 4) and the probability at each is given by the size of the jump in the distribution function:
$$
F_3 (x) = 
\begin{cases}
0, & x = 0\\
0.5, & 0 \leq x < 1\\
0.75, & 1 \leq x < 2\\
0.87, & 2 \leq x < 3\\
0.95, & 3 \leq x < 4\\
0.1, & 4 \leq x\\
\end{cases}
$$

drow plot of Distribution, Density, Survial, Hazard rate function of this Random Variable.

## Calculating Distribution function
\label{Dis}

According to the last slide we now that our cummulative distribution function is:

\begin{equation}
F_3 (x) = 
\begin{cases}
0, & x = 0\\
0.5, & 0 \leq x < 1\\
0.75, & 1 \leq x < 2\\
0.87, & 2 \leq x < 3\\
0.95, & 3 \leq x < 4\\
0.1, & 4 \leq x\\
\end{cases}
\end{equation}


## ploting Cummulative distribution function of  Model 3 
\label{CDF}
\tiny
```{r , warning=FALSE , echo=TRUE}
F_x3 = function(x){
  if(x < 0){
    prob = 0
  }
  if((x<1)&(x>=0)){
    prob = 0.5
  }
  if((x<2)&(x>=1)){
    prob = 0.75
  }
  if((x<3)&(x>=2)){
    prob = 0.87
  }
  if((x<4)&(x>=3)){
    prob = 0.95
  }
  if(x>=4){
    prob = 1
  }
  return(prob)
}
#please enter your number (input) as x:
x = 3
F_x3(x)
```
\normalsize


## Ploting Cummulative distribution function of Model 3
\tiny
```{r , echo=TRUE , out.height="55%" , out.width="85%" ,warning=FALSE}
S = seq(-10 , 10);w = c();k = 1
for(i in S){
  w[k] = F_x3(i)
  k = k+1}
plot(x = S , y = w , col = "blue" , type = "b" ,
     xlab = "x value" , ylab = "cummulative Probability",
     main = "Cummulative distribution function of Model 3")
```
\normalsize


## calculating Survial function of Model 3
\label{Sur}

According to bottem formula for calculating survial function we have:
\begin{equation}
S_X (x) = 1 - F_X (x)
\end{equation}
So our result is:
\begin{equation}
S_X (x) = 
\begin{cases}
0.5, & 0 \leq x < 1\\
0.25, & 1 \leq x < 2\\
0.13, & 2 \leq x < 3\\
0.05, & 3 \leq x < 4\\
0, & 4 \leq x\\
\end{cases}
\end{equation}

## Ploting Survial function of Model 3
\label{Psur}

```{r , echo=TRUE  , warning=FALSE}
S_x3 = function(x){
  prob = 1 - F_x3(x)
  return(prob)}
x = 3
S_x3(x)
```


## Ploting Survial function of Model 3
\small
```{r , echo=TRUE , out.height="55%" , out.width="85%" ,warning=FALSE}
S = seq(-10 , 10);q = c();k = 1
for(i in S){
  q[k] = S_x3(i)
  k = k+1}
 plot(x = S , y = q , col = "blue" , type = "b" ,
      xlab = "x value" , ylab = "Probability",
      main = "")
```
\normalsize

## calculating probability mass function of Model 3
\label{pdf}
Now here we want to Calculating the Probability mass function of Model 3 according to the bottem formula:
\begin{equation}
p_X (x)= Pr(X=x)
\end{equation}
\newline
So our result is:
\begin{equation}
p_3 (x) =
\begin{cases}
0.50, &x=0\\
0.25, & x=1\\
0.12, & x=2\\
0.08, & x=3\\
0.05, & x=4\\
\end{cases}
\end{equation}

## ploting probability mass function of Model 3
\label{Ppdf}

```{r , echo=TRUE , warning=FALSE}
 f_x3 = function(x){
   prob = F_x3(x) - F_x3(x-1)
   return(prob)
 }
x = 3
f_x3(x)
```


## ploting probability mass function of Model 3

```{r , echo=TRUE , out.height="55%" , out.width="85%" ,warning=FALSE}
S = seq(-10 , 10 );e = c();k = 1
for(i in S){
  e[k] = f_x3(i)
  k = k+1}
barplot( e  ,names.arg = -10:10 ,
         xlab = "X value" , ylab = "Probability" ,
         col = 5:1 , main = "Probability(mass) function")
```

## End
\huge   
- \textcolor{blue}{Thanks for your attention}